\documentclass[10pt, a4paper]{article}

\usepackage[osf]{libertine}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\linespread{1.05}
\selectfont


\usepackage{geometry}
\geometry{a4paper, textwidth=6.2in, textheight=9.5in, marginparsep=12pt, marginparwidth=1.4in}
% \setlength\parindent{0in}

\usepackage[usenames,dvipsnames]{xcolor}

\usepackage{marginnote}
\newcommand{\years}[1]{\marginnote{#1}}
\setlength{\marginparsep}{12pt}
\reversemarginpar

\usepackage{sectsty}
\usepackage[normalem]{ulem}
\sectionfont{\mdseries\upshape\Large}
\subsectionfont{\mdseries\scshape\normalsize}
\subsubsectionfont{\mdseries\upshape\large}

\usepackage[bookmarks, colorlinks, breaklinks]{hyperref}
\hypersetup{linkcolor=blue,citecolor=blue,filecolor=black,urlcolor=MidnightBlue}

\usepackage{graphicx}

\pagenumbering{gobble}


\begin{document}

\section*{\textsc{{\huge Interplay between Complex Orders in Functional Materials}\\[0.1cm] Didactic and scientific activities carried out during the Ph.D. course\\[0.2cm] {\normalsize Louis Ponet}}}

\vspace{0.4cm}
{\bf Courses}
\begin{itemize}
\item
Theory of Many-Body Systems.
\item
Seminars on Condensed Matter Physics.
\item
Quantum Information I.
\end{itemize}
\vspace{0.4cm}
{\bf Summer Schools, Workshops and Conferences}
\begin{itemize}
\item
``\textit{DPG Spring Meeting}''.\\
Dresden, Germany, 19-24 March 2017.\\
Contributed Talk.
\item
``\textit{SUCCESS Summer School}''.\\
Les Houches, France, 4-15 September 2017.
\item
``\textit{Nordita Winter School in Condensed Matter Physics}''.\\
Stockholm, Sweden, 21-31 Januari 2018.
\item 
``\textit{DPG Spring Meeting}''.\\
Berlin, Germany, 11-16 March 2018.\\
Contributed Talk.
\item
``\textit{Autumn School on Correlated Electrons}''. \\
J\"ulich, Germany, 17-21 September 2018.
\item
``\textit{8th MaNEP Winter School}''.\\
 Saas-Fee, Switzerland, 13-18 January 2019.
\item 
``\textit{DPG Spring Meeting}''.\\
Regensburg, Germany, 31 March-5 April 2019.\\
Contributed Talk.
\item
``\textit{TRR80: From Electronic Correlations to Functionality}''.\\
 Chiemsee, Germany, 24-28 June 2019.
\item
``\textit{Advanced Electronic Structure Methods in Condensed Matter Physics}''.\\
 Lausanne, Switzerland, 8-10 July 2019.
\item
``\textit{2nd School on Second Principles Density Functional Theory and SCALE-UP}''.\\
 Santander, Spain, 22-26 July 2019.
\item
``\textit{JEMS 2020}''.\\
 Lisbon, Portugal, 7-11 December 2020.\\
 Contributed Talk.
 \\
\item
``\textit{Ferro 2021}''.\\
 Online, 1-3 February 2021.\\
 Contributed Talk.
 \\

\end{itemize}
\vspace{0.4cm}
{\bf Scientific Visits}
\begin{itemize}
 \item
 Group of Prof. David Vanderbilt, Rutgers University, New-Brunswick, NJ, United States of America. \\
 25/05/2018 - 1/07/2018.
 \item
 Group of Prof. Stefan Bl\"ugel, Forschzungszentrun J\"ulich, J\"ulich, Germany.\\
 04/11/2018 - 09/11/2018.
\end{itemize}
\vspace{0.4cm}
\end{document}
