\documentclass[10pt, a4paper]{article}

\usepackage[osf]{libertine}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\linespread{1.05}
\selectfont


\usepackage{geometry}
\geometry{a4paper, textwidth=6.2in, textheight=9.5in, marginparsep=12pt, marginparwidth=1.4in}
% \setlength\parindent{0in}

\usepackage[usenames,dvipsnames]{xcolor}

\usepackage{marginnote}
\newcommand{\years}[1]{\marginnote{#1}}
\setlength{\marginparsep}{12pt}
\reversemarginpar

\usepackage{sectsty}
\usepackage[normalem]{ulem}
\sectionfont{\mdseries\upshape\Large}
\subsectionfont{\mdseries\scshape\normalsize}
\subsubsectionfont{\mdseries\upshape\large}

\usepackage[bookmarks, colorlinks, breaklinks]{hyperref}
\hypersetup{linkcolor=blue,citecolor=blue,filecolor=black,urlcolor=MidnightBlue}

\usepackage{graphicx}

\usepackage[final]{pdfpages}

\pagenumbering{gobble}

\begin{document}

\section*{\textsc{{\huge Interplay between Complex Orders in Functional Materials}\\[0.1cm] List of publications\\[0.2cm] {\normalsize Louis Ponet}}}

\hspace{2cm}

\begin{itemize}
	\item[$\bullet$]
	{\bf Louis Ponet}, S. Artyukhin\\
	{\it First-principles theory of giant Rashba-like spin splitting in bulk GeTe}\\
	\href{https://doi.org/10.1103/PhysRevB.98.174102}{Phys. Rev. B. {\bf 98}, 174102 (2018)}
	\item[$\bullet$]
	P. Chen, {\bf Louis Ponet}, K. Lai, R. Cingolani, S. Artyukhin\\
	{\it Domain wall-localized phonons in BiFeO$_3$: spectrum and selection rules}\\
	\href{https://doi.org/10.1038/s41524-020-0304-y}{npj Comput. Mater. {\bf 6}, 48 (2020)}
	\item[$\bullet$]
	Y. Huang, L. Zheng, P. Chen, X. Cheng, S. Hsu, T. Yang, X. Wu, {\bf Louis Ponet}, R. Ramesh, L. Chen, S. Artyukhin, Y. Chu, K. Lai\\
	{\it Unexpected giant microwave conductivity in a nominally silent BiFeO$_3$ domain wall}\\
	\href{https://doi.org/10.1002/adma.201905132}{Adv. Mater. {\bf 32}, 1905132 (2020)}
	\item[$\bullet$]
	C. Stefani, {\bf Louis Ponet}, K. Shapovalov, P. Chen, E. Langenberg, D. G. Schlom, S. Artyukhin, M. Stengel, N. Domingo, G. Catalan\\
	{\it Mechanical softness of ferroelectric 180$^\circ$ domain walls}\\ 
	\href{https://doi.org/10.1103/PhysRevX.10.041001}{Phys. Rev. X {\bf 10}, 041001 (2020)}
	\item[$\bullet$]
	O. Yu. Gorobtsov, {\bf Louis Ponet}, S. K. K. Patel, N. Hua, A. G. Shabalin, S. Hrkac, J. Wingert, D. Cela, J. M. Glownia, D. Zhu, R. Medapalli, M. Chollet, E. E. Fullerton, S. Artyukhin, O. G. Shpyrko, A. Singer\\
	{\it Femtosecond control of phonon dynamics near a magnetic order critical point}\\
	Nature communications {\it under review}
	\item[$\bullet$]
	F. Fan, Z. Gao, {\bf Louis Ponet}, J. Wang, H. Huang, S. Artyukhin, J. Hong, S. Cheong, X. Wang\\
	{\it Visualization of large-scale charged domain wall in hexagonal manganites}\\
	Applied Physics Letters {\it accepted}
	\item[$\bullet$]
	{\bf Louis Ponet}, S. Artyukhin, Th. Kain, Anna Pimenov, A. Shuvaev, X. Wang, S. Cheong, M. Mostovoy, Andrei Pimenov\\
	{\it Magnetoelectric Crankshaft}\\
	Nature {\it under review}
	\item[$\bullet$]
	B. Dhanabalan, G. Biffi, A. Moliterni, V. Olieric, C. Giannini, G. Saleh, {\bf Louis Ponet}, M. Prato, M. Imran, L. Manna, R. Krahne, S. Artyukhin, M. P. Arciniegas\\
	{\it Engineering the optical emission and robustness of metal-halide layered perovskites through ligand accommodation}\\
	Advanced Materials {\it accepted}
\end{itemize}% \includepdf[pages=-]{DissipationAQC.pdf}
% \includepdf[pages=-]{RSVD.pdf}
% \includepdf[pages=-]{EntropyProdCM.pdf}
% \includepdf[pages=-]{ExtractableWorkQB.pdf}
% \includepdf[pages=-]{PersistentCurrents.pdf}

\end{document}

